Conflict Analysis for Mixed-integer SDPs

### 2022-08-01 tex/ConflictAnalysis.tex
add question to tex document
add explanation about primal solutions

A primal ray, i.e., a point $(\hat{X}, \hat{z}, \hat{r}^\ell, \hat{r}^u)$
satisfying the system
\begin{equation}\label{eq:ray}
\begin{aligned}
  & \skal{A^j}{X} + (D\T z)_j + r^\ell_j - r^u_j = 0 && \forall\; j \in [m],\\
  & \skal{A^0}{X} + d\T z + \ell\T r^\ell - u\T r^u > 0,
\end{aligned}
\end{equation}
yields an inequality as in~\eqref{eq:ConflictCut}
or~\eqref{eq:ConflictCut2}.

@@

>>>>

A primal ray, i.e., a point $(\hat{X}, \hat{z}, \hat{r}^\ell, \hat{r}^u)$
satisfying the system
\begin{equation}\label{eq:ray}
\begin{aligned}
  & \skal{A^j}{X} + (D\T z)_j + r^\ell_j - r^u_j = 0 && \forall\; j \in [m],\\
  & \skal{A^0}{X} + d\T z + \ell\T r^\ell - u\T r^u > 0,
\end{aligned}
\end{equation}
\todo[inline]{Frederic: I think we also need~$X \succeq 0$ and~$z,\, r^{\ell},\, u
  \geq 0$ in~\eqref{eq:ray}.}
yields an inequality as in~\eqref{eq:ConflictCut}
or~\eqref{eq:ConflictCut2}.

@@

% -------------------------------------------------------------------------
\subsection{Primal Solutions for Special Cases}
\label{sec:Primal}

In order to produce conflict constraints, one needs solutions for the
primal of the LP-relaxation~\eqref{SDP-P}. Because SCIP-SDP preprocesses
certain SDP problems, we briefly discuss how primal solutions can be
obtained in these cases.

% -------------------------------------------------------------------------
\subsubsection{SDPs with One Variable}

For the special case of SDP-relaxations in which only one variable is
unfixed, SCIP-SDP uses specialized methods to solve them. In order to
derive a primal solution, consider
\begin{equation}\label{eq:basic}
  \inf\; \{\gamma\, \mu \suchthat \mu\, A - B \succeq 0,\; \ell \leq \mu \leq u\},
\end{equation}
where $A$, $B \in \R^{n \times n}$ are symmetric matrices, $\gamma \in \R$,
and $\ell \in \R \cup \{-\infty\}$, $u \in \R \cup \{\infty\}$.

The dual SDP of~\eqref{eq:basic} is
\begin{equation}\label{eq:dual}
  \sup\; \{\skal{B}{X} + \ell\, \alpha - u\, \beta \suchthat \skal{A}{X} +
  \alpha - \beta = \gamma,\; X \succeq 0,\; \alpha,\; \beta \geq 0\}.
\end{equation}

If $\mu^*$ is an optimal solution for~\eqref{eq:basic}, we can construct an
optimal primal solution $(X^*, \alpha^*, \beta^*)$ as follows.

\paragraph{Case 0, $\boldsymbol{\gamma = 0}$:}

Here we can take $(X^*, \alpha^*, \beta^*) = (0,0,0)$. Thus, in the
following, we assume that $\gamma \neq 0$.

\paragraph{Case 1, $\boldsymbol{\ell < \mu^* < u}$:}

Then by complementary slackness necessarily $\alpha^* = \beta^* = 0$, and
thus $\skal{A}{X^*} = \gamma$. Optimality implies that
\[
\gamma\, \mu^* = \skal{B}{X^*}
\quad\Leftrightarrow\quad
\mu^* \skal{A}{X^*} = \skal{B}{X^*}
\quad\Leftrightarrow\quad
\skal{\mu^* A - B}{X^*} = 0.
\]

Note that in~\eqref{eq:dual} there is only one affine constraint, which
implies that there exists a rank-1 solution. Thus, we are aiming to
construct a rank-1 solution $X^* = x^* (x^*)\T \succeq 0$.  This solution
has to satisfy:
\[
0 = \skal{\mu^* A - B}{x^* (x^*)\T} = (x^*)\T (\mu^* A - B) x^*.
\]
Therefore, $x^*$ needs to be an eigenvector for the eigenvalue~$0$ of
$\mu^* A - B$ and thus $x^* \in \ker(\mu^* A - B)$. Moreover,
$\skal{A}{x^*(x^*)\T} = (x^*)\T A x^* = \gamma$ needs to hold.  Consider a
generator matrix $C \in \R^{n \times r}$ of the kernel of $\mu^* A - B$,
i.e., $\ker(\mu^* A - B) = \{C y \suchthat y \in \R^r\}$.  Then by scaling,
the problem reduces to the question whether there exists $y \in \R^r$ such
that
\[
y\T C\T A C y
\begin{cases}
  > 0 & \text{if } \gamma > 0,\\
  < 0 & \text{if } \gamma < 0.
\end{cases}
\]
Computing a maximal/minimal eigenvalue/eigenvector of $C\T A C$ will decide
this.

\paragraph{Case 2, $\boldsymbol{\mu^* = \ell}$:}

In this case, $\beta^* = 0$ has to hold.

If $\gamma \geq 0$, we can take
$(X^*, \alpha^*, \beta^*) = (0, \gamma, 0)$, which is feasible with
objective value $\gamma \ell = \gamma \mu^*$ as needed. Therefore, this
point is optimal.

If $\gamma < 0$, the conditions are
\[
\skal{B}{X^*} + \ell\, \alpha^* = \gamma\, \ell,\quad
\skal{A}{X^*} + \alpha^* = \gamma,\quad X^* \succeq 0,\; \alpha^* \geq 0.
\]
This yields
\[
\skal{B}{X^*} + \ell\, \alpha^* = \skal{A}{X^*}\, \ell + \alpha^*\, \ell
\quad\Leftrightarrow\quad
\skal{\ell\, A - B}{X^*} = 0.
\]
As above, we are looking for a rank-1 solution $(x^*)\T x^*$. Let
$C \in \R^{n \times r}$ be a generator matrix of $\ker(\ell\, A - B)$. Then
we need $y^* \in \R^r$ such that
\[
(y^*)\T C\T A C y^* < 0.
\]
If this is impossible, the problem is infeasible (e.g., if $A \succeq
0$). Otherwise, we can scale $x^* = C y^*$ and set $\alpha^* = 0$.

\paragraph{Case 3, $\boldsymbol{\mu^* = u}$:}

This case is symmetric to Case 2. If $\gamma < 0$, we can set
$(X^*, \alpha^*, \beta^*) = (0, 0, -\gamma)$.

If $\gamma \geq 0$, we can construct a rank-1 solution as above.
